---
title: "Data structures"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Opening quiz:

1. What are the three properties of a vector, other than its contents?
	* Length, type, and...attributes? (like names?)
2. What are the four common types of atomic vectors? What are the two rare types?
	* boolean, numeric (integer and double), factor, string. raw and complex?
3. What are attributes? How do you get them and set them?
	* Not really sure about either.
4. How is a list different from an atomic vector? How is a matrix different from a data frame?
	* A list is different than an atomic vector in that it can contain heterogeneous types. It can also be nested. Atomic vectors are homogeneous and flat. A matrix is different than a dataframe in that it is homogenous, like a vector. A dataframe can contain heterogenous types.
5. Can you have a list that is a matrix? Can a data frame have a column that is a matrix?
	* I don't think you can have a list that *is* a matrix, but I think you can have a list of length one that *contains* a matrix?
	* I bet a data frame could have a column that is a matrix, because it can have columns that are vectors, lists, or data frames. But I've never seen a data frame with a column that is a matrix.

## Vectors

A list *is* a vector. But an atomic vector is not a list. So, hierarchically:

* Vectors
	* Atomic vectors
	* Lists (non-atomic vectors?)

* `is.vector()` doesn't test whether an object is a vector; it returns `TRUE` only if the object is a vector with no attributes apart from names. 
	* Q: why?
* `c()` is short for "combine"!

* Lists are sometimes called *recursive* vectors because lists can contain other lists.
	* Q: is *recursive* meant to be the opposite of *atomic*?

### Exercises

1. What are the six types of atomic vector? How does a list differ from an atomic vector?
	* boolean
	* integer
	* double
	* character
	* complex
	* raw
	* A list is different than an atomic vector in that it can contain heterogeneous types and it can contain nested structures. (It's sometimes called a *recursive* vector.) An atomic vector is always flat and all the values it contains must have the same type. 

2. What makes `is.vector()` and `is.numeric()` fundamentally different to `is.list()` and `is.character()`?
  * What does `is.vector()` do again? I guess it just checks whether an object is a vector that doesn't have attributes other than names? I'm not sure how to link that up with `is.numeric()`. I know that `is.list()` only returns true on... lists, i.e. recursive, i.e. non-atomic vectors. But `is.numeric()` and `is.character()` only return true for atomic vectors. So I'm confused about this one. Come back to it.

3. Test your knowledge of vector coercion rules by predicting the output of the following uses of c():

```
c(1, FALSE) # a double vector, 1 and 0
c("a", 1) # a character vector, "a" and "1"
c(list(1), "a") # a character vector "1" and "a"
c(TRUE, 1L) # an integer vector, 1L and 1L
```

```{r}
c(1, FALSE) # a double vector, 1 and 0 CORRECT
c("a", 1) # a character vector, "a" and "1" CORRECT
c(list(1), "a") # a character vector "1" and "a" INCORRECT
c(TRUE, 1L) # an integer vector, 1L and 1L CORRECT
```

> I got that third one wrong! I was thinking that `c()` would coerce everything to vectors, because I think I still think of `c()` as "create a vector", but in fact it just means `combine()`. So it first coerced the `"a"` to a list, and then it created a list of two lists.

4. Why do you need to use unlist() to convert a list to an atomic vector? Why doesn’t as.vector() work?
  * Because a list is already a vector, just not an atomic one.

5. Why is 1 == "1" true? Why is -1 < FALSE true? Why is "one" < 2 false?
  * In the first and second cases, the inputs on either side of the operator can be coerced to the same type, namely double. But in the third case, `"one"` can't be coerced to a numeric.

6. Why is the default missing value, NA, a logical vector? What’s special about logical vectors? (Hint: think about c(FALSE, NA_character_).)
  * What's special about logical vectors is that they can be coerced all the way from boolean up to character. `FALSE` can become `1L`, which can become `1`, which can become `"1"`. So if you have a vector of any type and you put `NA` in it, that can always be coerced to the type you're looking for. (And conversely, that default logical NA isn't going to coerce *other* inputs into logicals.)
  * If you do `c(FALSE, NA_character_)`, you should end up with `"FALSE"` and `NA_character_`

## Attributes

* You can remove names from a named vector with `unname()`.
* A factor vector is just an integer vector with two special attributes:
  * A `class`, "factor", which makes them behave differently than regular integer vectors. (Maybe you can't use arithmetic on them? Correct! `factor(c("a", "b", "c")) + 1` returns `‘+’ not meaningful for factors`)
  * `levels`, which defines the set of allowed values.
* I didn't know you could use `table` on a factor vector

### Exercises

1. An early draft used this code to illustrate structure():

```
structure(1:5, comment = "my attribute")
## [1] 1 2 3 4 5
```

But when you print that object you don’t see the comment attribute. Why? Is the attribute missing, or is there something else special about it? (Hint: try using help.)

> The attribute is not missing. The comment attribute is special — it isn't printed by default. You can access it with `comment()`, or with `attr(my_vector, "comment")`.

2. What happens to a factor when you modify its levels?

```{r}
f1 <- factor(letters)
f1

levels(f1) <- rev(levels(f1))
f1

levels(f1)

f1
```

> The `levels` attribute changes, which changes the allowed/disallowed values, and also changes the ordering of those values.
> Oh wow. The values themselves changed order too... I'm confused about this.
> Q: why did that happen?

3. What does this code do? How do f2 and f3 differ from f1?

```{r}
f2 <- rev(factor(letters))
f2

f3 <- factor(letters, levels = rev(letters))
f3
```

> f2 is different than f1 because it just reverses the order of the values in the vector without reversing the order of the levels.
> f3 is the opposite — it reverses the order of the levels without reversing the order of the vector values.