---
title: "Environments"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quiz

1. List at least three ways that an environment is different to a list.

> I have no idea how environments are represented. I don't really understand the context of the question, because environments and lists seem entirely different.

2. What is the parent of the global environment? What is the only environment that doesn’t have a parent?

> Loaded packages?

3. What is the enclosing environment of a function? Why is it important?

> The environment in which the will be evaluated? I assume it's important because it determines how the scoping rules interact when a function is called.

4. How do you determine the environment from which a function was called?

> No idea.

5. How are <- and <<- different?

> I've seen this before, but I can't remember.

## Prerequisites

```{r}
library(rlang)

# Some API changes that haven't made it in rlang yet
search_envs <- function() {
  rlang:::new_environments(c(
    list(global_env()),
    head(env_parents(global_env()), -1)
  ))
}
```

## Environment basics

K, before we get into this, let's get this clear: an environment is a data structure. That makes sense. You can define its contents with `rlang::env()`, which works like `list()`.

For example:

```{r}
e1 <- env(
  a = FALSE,
  b = "a",
  c = 2.3,
  d = 1:3,
)
```

Notes:

- The job of an environment is to associate, or bind, a set of names to a set of values.
- The objects in an environment have no inherent order.
- An environment can contain itself.
- To print an environment, use `rlang::env_names()`.

### Important environments

- the global environment, the current environment, the empty environment (which is the only one that doesn't have a parent)
- not sure why this matters yet. Clearly it's relevant for lexical scoping, but I'm not sure when it would be relevant in situations where you have carefully named all your variables to avoid unintended name masking...

### Super assignment

`<<-` doesn't define a variable in the current environment; it redefines a variable in the parent environment. E.g.

```{r}
x <- 0
f <- function() {
  x <<- 1
}
f()
x
```

### Packages and environments

Each attached package becomes the immediate parent of the global environment. The immediate parent of the global environment is whichever was the last package you attached. The parent of that environment is the previous package you attached, etc. This is where lexical scoping and environments interact to actually matter for day-to-day purposes — especially relevant when two packages have functions with the same name. Though in package development that shouldn't matter, because you'd always prepend a function call with the package name.

