---
title: "Functions"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quiz

1. What are the three components of a function?
> its name, its arguments, and its contents?
> nope: body, arguments, and environment

2. What does the following code return?

```
x <- 10
f1 <- function(x) {
  function() {
    x + 10
  }
}
f1(1)()
```

> 20? Let's see...

```{r}
x <- 10
f1 <- function(x) {
  function() {
    x + 10
  }
}
f1(1)()
```

> Lol. Fooled me.

3. How would you more typically write this code?

```{r}
`+`(1, `*`(2, 3))
```

Like this:

```{r}
1 + 2*3
```

> correct

4. How could you make this call easier to read?

```{r}
mean(, TRUE, x = c(1:10, NA))
```

Like this:

```{r}
mean(x = c(1:10, NA), na.rm = TRUE)
```

> correct

5. Does the following function throw an error when called? Why/why not?

```
f2 <- function(a, b) {
  a * 10
}
f2(10, stop("This is an error!"))
```

> I expect that if we passed in something like b = 5, we'd get an error because b never got used in the function. But since we're passing in an error message, maybe that nullifies the "never got used" error because a different error was passed in instead? But it's not like it was actually used, so that wouldn't really make sense. I have no idea. Let's see.

```{r}
f2 <- function(a, b) {
  a * 10
}
f2(10, stop("This is an error!"))
```

> Okay, turns out I was wrong about that initial claim too. The function runs fine even if you give it a normal input. (I wonder why I've seen that error before in other situations?)

6. What is an infix function? How do you write it? What’s a replacement function? How do you write it?

> No clue.

7. What function do you use to ensure that a cleanup action occurs regardless of how a function terminates?

> No clue.
> answer: `on.exit()`. That could be very handy!

## Function components

### Exercises

1. What function allows you to tell if an object is a function? What function allows you to tell if a function is a primitive function?

> `is.function`
> `is.primitive`

2. This code makes a list of all functions in the base package.

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
```

Use it to answer the following questions:

  a. Which base function has the most arguments?
  b. How many base functions have no arguments? What’s special about those functions?
  c. How could you adapt the code to find all primitive functions?

```{r}
arg_lengths <- sapply(funs, function(x) length(formals(x)))

# a: `scan`
names(which.max(arg_lengths))

# b: 224 have no arguments. Most of them are primitive...but not all?
no_arg_functions <- names(arg_lengths[arg_lengths == 0])

length(no_arg_functions)

mean(sapply(funs[no_arg_functions], is.primitive))

# c: change `is.function` to `is.primitive`, presumably?
objs <- mget(ls("package:base"), inherits = TRUE)
prim_funs <- Filter(is.primitive, objs)

length(prim_funs)

length(funs)
```


3.What are the three important components of a function?
> Its formals (arguments), its body (the code in the body of the function), and its environment.

4. When does printing a function not show what environment it was created in?
> When it was created in the global environment?

Q: need more information about environments. I don't have a strong enough understanding of this.

## Lexical scoping

Scoping: the set of rules that defines how a value is extracted from a symbol.

There are two types of scoping in R: "lexical" scoping and "dynamic" scoping. Not going to worry about the difference for now, except to say that we're learning about lexical scoping at the moment.

"lexical" comes from the computer science term "lexing" which is the process of converting code represented as text to meaningful objects that the computer can manipulate. Seems kind of like the converse of the definition of scoping...

Different languages implement scoping in different ways. R's implementation relies on four concepts:

- name masking
- functions vs variables
- a fresh start
- dynamic lookup

### Name masking

```{r}
f <- function() {
  x <- 1
  y <- 2
  c(x, y)
}
f()
rm(f)
```
```{r}
x <- 2
g <- function() {
  y <- 1
  c(x, y)
}
g()
rm(x, g)
```

```{r}
x <- 1
h <- function() {
  y <- 2
  i <- function() {
    z <- 3
    c(x, y, z)
  }
  i()
}
h()
rm(x, h)
```

When a name isn't defined in a given function, R will look one level up — in the exterior function (if there is one), then in the global environment, then in other loaded packages.

A *closure* is a function created by another function. E.g. in the code below, `k` is a closure.

```{r}
j <- function(x) {
  y <- 2
  function() {
    c(x, y)
  }
}
k <- j(1)
k()
rm(j, k)
```

`j` calls `function() {c(x, y)}`, so you might wonder how it knows the value of `y`. It knows because closures preserve the environments they were created in. In this case, that environment...'is'...the function `j`?

Q: this all makes sense, but why is it called "name masking"?

### Functions vs. variables

The rules for finding functions are the same as the rules for finding variables - look first in the immediate environment, then up, then up, etc. But there is one tweak: if a name is used in a context where it's obvious that you're referring to a function, R will call the function even if a variable would otherwise have masked that name.

```{r}
n <- function(x) x / 2
o <- function() {
  n <- 10
  n(n)
}
o()
rm(n, o)
```

### A fresh start

In between executions of a function, any names that were created during the first execution will be wiped clean before the second. For example:

```{r}
j <- function() {
  if (!exists("a")) {
    a <- 1
  } else {
    a <- a + 1
  }
  a
}
j()
rm(j)
```

The function above returns 1 every time. This is because every time a function is called, R creates a new environment to host the execution. Every execution is independent.

### Dynamic lookup

R looks up the value of a name in a function when the function is executed, not when it's created.

```{r}
f <- function() x
x <- 15
f()
```

```{r}
x <- 20
f()
```

(This function is not referentially transparent!) You can use `codetools::findGlobals()` to find all the external dependencies of a function.

#### Exercises

1. What does the following code return? Why? What does each of the three c’s mean?

```{r}
c <- 10
c(c = c)
```

> It returns a numeric vector of length one with the only element having the name "c" and the value 10.
> In the second line of code, the first c is a function — short for "combine", used to create a vector — the second c is a name given to the first element in the vector, and the third c refers to the variable "c" which was defined in the line above.

2. What are the four principles that govern how R looks for values?

- name masking: a name can mask the value of another name if it is defined in a more immediate environment relative to the execution of a function.
- functions vs variables: functions follow the same rules as variables except that when you use the name of a function in a way where it's obvious that you're referring to a function (e.g. `f()` rather than `f`), R will look for the matching *function* that is most immediate, even if that means ignoring a more immediate variable.
- "a fresh start": every time you execute a function, any names that are defined internally will be redefined as if for the first time, even if that function was executed previously.
- dynamic lookup: R looks for the value of a variable when the function is executed, not when it's defined. This means that if you source a function that has undefined variables, you won't get an error — you would need to execute the function for R to actually search for those variables and notice that they're undefined.

3. What does the following function return? Make a prediction before running the code yourself.

```{r}
f <- function(x) {
  f <- function(x) {
    f <- function(x) {
      x ^ 2
    }
    f(x) + 1
  }
  f(x) * 2
}
f(10)
```

1. the first internal f is defined.
2. the first internal f is called with 10 as its argument.
3. the second internal f is defined.
4. the second internal f is called, with 10 as its argument, returning 100, then 1 is added to the result, returning 101.
5. 101 is multiplied by 2, returning 202.

Prediction: 202

## Every operation is a function call

> “To understand computations in R, two slogans are helpful:
> - Everything that exists is an object.
> - Everything that happens is a function call."
> — John Chambers

Everything that happens in R is a function call, even if it doesn't look like one. This is true of for loops, conditionals, subsetting, and mathematica operations. This can be useful in conjunction with sapply, vapply, etc. For example:

```{r}
x <- list(1:3, 4:9, 10:12)
sapply(x, "[", 2)
```

## Function arguments

There is a difference between an argument's formal arguments (which are a property of a function) and its calling arguments, which are the values passed in to those formal arguments when the function is executed.

### Calling functions

You can abbreviate the names of the formal arguments when you execute the function... but I have no clue why you would do that. This seems like the kind of thing that gives R a bad reputation.

### Calling a function given a list of arguments

If you have a list of arguments and you want to pass them into a function for execution, you can use the function `do.call`. For example:

```{r}
args <- list(1:10, na.rm = TRUE)
do.call(mean, args)

# Equivalent to
mean(1:10, na.rm = TRUE)
```

### Default and missing arguments

You can apparently define default arguments in terms of variables that get defined within that seame function... Also seems kinda crazy.

You can use the function `missing()` to determine whether an argument was supplied or not.

### Lazy evaluation

In R, function arguments are only evaluated if they're actually used.

```{r}
f <- function(x) {
  10
}
f(stop("This is an error!"))
```

If you want to force an argument to be evaluated, you can use `force()`.

```{r}
f <- function(x) {
  force(x)
  10
}
f(stop("This is an error!"))
```

```{r}
add <- function(x) {
  function(y) x + y
}
adders <- lapply(1:10, add)
adders[[1]](1)
adders[[10]](10)
```

```{r}
add <- function(x) {
  force(x)
  function(y) x + y
}
adders2 <- lapply(1:10, add)
adders2[[1]](1)
adders2[[10]](10)
```

```{r}
add <- function(x) {
  x
  function(y) x + y
}
adders3 <- lapply(1:10, add)
adders3[[1]](1)
adders3[[10]](10)
```


```{r}
adders[[1]](1)
```

These examples don't make sense to me. Looks like it might be an error — perhaps I should email Hadley about this, and also my question about name masking.

Another useful trick:

Use `&&` or `||` to evaluate the two conditions in a predicate in turn. Especially useful when you want to check whether something is null before checking whether the conditionally non-null value is in some given range. E.g

```{r}
x <- NULL
if (!is.null(x) && x > 0) {
  print("hey")
} else {
  print("ho")
}
```

#### Exercises

1. Clarify the following list of odd function calls:

```{r}
set.seed(1)
x <- sample(replace = TRUE, 20, x = c(1:10, NA))
y <- runif(min = 0, max = 1, 20)
cor(m = "k", y = y, u = "p", x = x)
```

```{r}
set.seed(1)
x <- sample(x = c(1:10, NA), size = 20, replace = TRUE)
y <- runif(n = 20, min = 0, max = 1)
cor(x = x, y = y, method = "kendall", use = "pairwise.complete.obs")
```

2. What does this function return? Why? Which principle does it illustrate?

```{r}
f1 <- function(x = {y <- 1; 2}, y = 0) {
  x + y
}
f1()
```

It returns 3 because when x is evaluated, a new y is defined in the local environment which masks the original value of y, 0, and then x is set to 2. This illustrates name masking.

3. What does this function return? Why? Which principle does it illustrate?

```{r}
f2 <- function(x = z) {
  z <- 100
  x
}
f2()
```

This function returns 100, because argument defaults can be defined in terms of variables that don't get defined until the body of the function. (What if `z` existed outside this function?) This illustrates the principle of lazy evaluation.

## Special calls

### Infix functions

You can write user-defined functions that are called in between arguments instead of outside of them, the same way these functions are equivalent:

```{r}
1 + 2
`+`(1, 2)
```

The magrittr pipe is an example of this!

```{r}
tibble::tibble(a = c(1)) %>% mutate(b = 2)
`%>%`(`%>%`(tibble::tibble(a = c(1)), mutate(b = 2)), select(a))
```

### Replacement functions

You can write functions with a special name format, namely `xxx<-`, that act like they modify their arguments in place. (They don't actually — they create a modified copy of the data.) But if you want to save yourself from using that extra memory, you should use built-in functions. Replacement and subsetting are useful in tandem, like in the example below:

```{r}
x <- c(a = 1, b = 2, c = 3)
names(x)
names(x)[2] <- "two"
names(x)
```

### Exercises

1. Create a list of all the replacement functions found in the base package. Which ones are primitive functions?

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
replacement_funs <- funs[str_detect(names(funs), ".*<-$")]
names(replacement_funs)

prim_replacement_funs <- Filter(is.primitive, replacement_funs)
names(prim_replacement_funs)
```

2. What are valid names for user-created infix functions?

> Anything that starts and ends with `%`.

3. Create an infix xor() operator.

```{r}
`%xor%` <- function(a, b) {
  if (a & !b | b & !a) return(TRUE) else return(FALSE)
}
FALSE %xor% FALSE
```

4. Create infix versions of the set functions intersect(), union(), and setdiff().

```{r}
a <- c(3, 4, 5, 6, 3, 3, 3, 1)
b <- c(1, 2, 3, 4, 4)

`%intersect%` <- function(a, b) {
  a_index <- a %in% b
  unique(a[a_index])
}
all(a %intersect% b == intersect(a, b))
all(b %intersect% a == intersect(b, a))

`%union%` <- function(a, b) {
  unique(c(a, b))
}
all(a %union% b == union(a, b))
all(b %union% a == union(b, a))

`%setdiff%` <- function(a, b) {
  a_index <- !a %in% b
  unique(a[a_index])
}
all(a %setdiff% b == setdiff(a, b))
all(b %setdiff% a == setdiff(b, a))

setdiff(c(1, 2, 3), c(3, 4, 5))
```

5. Create a replacement function that modifies a random location in a vector.

```{r}
`random_location<-` <- function(x, value) {
  random_index <- sample(1:length(x), 1)
  x[random_index] <- value
  return(x)
}
a <- c(1, 2, 3, 4)
random_location(a) <- 5
a
```

